Pure Spin Current Injection in Hydrogenated Graphene Structures

### 2017-07-20 mu/article/article.tex
Finishing corrections

Graphene, an allotrope of carbon with hexagonal 2D lattice structure presents
properties like fractional quantum Hall effect at room temperature, excellent
thermal transport properties, excellent conductivity\cite{heerscheNat07} and
strength \cite{geimNM07, reinaNL08, novoselov2S07, balandinNL08} being then a
perfect platform to be used in two-dimensions electronic systems; however most
electronic applications are disabled by the absence of a semiconducting gap.
Recent studies demonstrate that the band gap of graphene can be opened by
applying an electric field,\cite{zhangN09} reducing the surface
area,\cite{hanPRL07} or applying uniaxial strain.\cite{niACSN08} Another
possibility to open the gap is by doping; this has been successfully achieved
using nitrogen,\cite{weiNL2009} boron-nitrogen,\cite{guoIJ11}
silicon,\cite{colettiPRB10} noble-metals,\cite{varykhalovPRB10} and
hydrogen.\cite{eliasS09, guisingerNL09, samarakoonACSN10}
% 
Depending on the percentage of hydrogenation and spatial configurations of
hydrogen-carbon bonds, hydrogenated graphene can result in different spatial
configurations.
% 
In this paper we present two 50\% hydrogenated graphene noncentrosymmetric
structures both presenting a discernible band gap: the \emph{up} structure,
shown in Fig. \ref{fig:up-struc}, has hydrogen atoms bonded to the carbon layer
only in the upper side of the structure while the \emph{alt} structure, shown
in Fig. \ref{fig:alt-struc}, has hydrogen alternating in the upper and bottom
sides of the carbon slab.\cite{zapataPSB2016}

@@

This paper is organized as follows. In Section \ref{sec:theory} we present the
theory and formulas that describe PSC and SVI. In Section \ref{sec:results} we
describe the details of calculations and the corresponding SVI spectra for the
\emph{up} and \emph{alt} structures. Finally, we present our conclusions in
Section \ref{sec:conclusions}.

@@

\caption{Unit cell of \emph{up} structure. Layer division, atom types and
positions for the \emph{up} structure. The structure unit cell was divided in
two layers corresponding to hydrogen and carbon atoms.The corresponding layer
atom position is depicted in Fig. \ref{fig:up-struc} with the corresponding
number of layer.}
\label{tab:up-unitcell}
\end{table}
% 
% 
\begin{table}[t]
\center
\begin{tabular}{ccccc}\\
\hline
\quad Layer \quad & \quad Atom \qquad & \multicolumn{3}{c}{Position [\AA]} \\
\cline{3-5}
\quad No.   \quad & \quad type \qquad & $x$ & $y$ & $z$  \\
\hline
1 & H &  -0.61516 &  -1.42140 & \ 1.47237 \\
2 & C &  -0.61516 &  -1.73300 & \ 0.39631 \\
3 & C & \ 0.61516 & \ 1.73300 & \ 0.15807 \\
4 & C & \ 0.61516 & \ 0.42201 &  -0.15814 \\
5 & C &  -0.61516 &  -0.37396 &  -0.39632 \\
6 & H &  -0.61516 &  -0.68566 &  -1.47237 \\
\hline
\end{tabular}

\caption{Unit cell of \emph{alt} structure. Layer division, atom types and
positions for the \emph{alt} structure. The structure unit cell was divided in
six layers corresponding each one to atoms in different $z$ positions. The
corresponding layer atom position is depicted in Fig. \ref{fig:alt-struc} with
the corresponding number of layer.}
\label{tab:alt-unitcell}
\end{table}

We preset the results for 
and
$\mathcal{V}_{\sigma^{\mathrm{b}}}(\omega,\alpha)$ 
and
$\mathcal{V}_{\mathrm{a}}(\omega,\alpha)$ 
for the
C$_{16}$H$_{8}$-up  and C$_{16}$H$_{8}$-alt 
structures being both
noncentrosymmetric semi-infinite 2D carbon systems with 50\% hydrogenation in
different arrangements. We recall that the \emph{up} structure has hydrogen
atoms only on the upper side of the carbon sheet while the \emph{alt} structure
has alternating hydrogen atoms on the upper and bottom sides. Also we take the
hexagonal carbon lattice to be on the $xy$ plane for both structures, and the
carbon-hydrogen bonds on the perpendicular $xz$ plane, as depicted in Figs.
\ref{fig:up-struc} and \ref{fig:alt-struc}. The coordinates for the
\emph{up} and \emph{alt} unit cells of the structures are presented in Tables
\ref{tab:up-unitcell} and \ref{tab:alt-unitcell}.

We calculated the self-consistent ground state and the Kohn-Sham
states using density functional theory in the local density approximation (DFT-
LDA) with a planewave basis using the ABINIT code \cite{gonzeCPC09}.
% 
We used Hartwigsen-Goedecker-Hutter (HGH) relativistic separable dual-space
Gaussian pseudopotentials \cite{hartwigsenPRB98} including the spin-orbit
interaction needed to calculate $\mu^{\mathrm{abcd}}(\omega,\alpha)$ presented
in Eq. \eqref{eq:mu}.
% 
The convergence parameters for the calculations of our results corresponding to
the \emph{up} and \emph{alt} structures are cutoff energies of 65\,Ha and
40\,Ha, resulting in LDA energy band gaps of 0.088\,eV and 0.720\,eV,
respectively. The energy eigenvalues and matrix elements for the \emph{up} and
\emph{alt} structures were calculated using 12802 $\mathbf{k}$ points and 14452
$\mathbf{k}$ points in the IBZ to integrate $\mu^{\mathrm{abcd}}(\go)$ 
and $\xi^{\rma\rmb}(\go)$ using the linearized analytic tetrahedron
method (LATM).\cite{nastosPRB07}
We neglect the
anomalous velocity term $\hbar(\bfgs\times\nabla V)/4m^2c^2$, 
where $V$ is the crystal potential, in 
 $\hat\bfv$ of Eq.~\eqref{z.1}, as this term is known to give
small contribution to PSC.\cite{bhatPRL05}
Therefore, $[\hat\bfv,\hat\bfS]=0$ and 
Eq.~\eqref{z.1} reduces to $\hat K^{\rma\rmb}=\hat v^\rma \hat
S^\rmb=\hat S^\rmb \hat v^\rma$.

@@

\begin{table}[b]
\begin{tabular}{cccccc}
\hline
\multirow{2}{*}{Structure \quad} & 
Kind of \quad & 
Pol. &
Energy & 
\multicolumn{2}{c}{$\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)$}\\
\cline{5-6}
& system & Ang. & [eV] & $\mathrm{ab}$ \quad & [Km/s]\\
\hline
\emph{up}    & 2D   & 40    & 0.088  & $\mathrm{yz}$ &  87.2    \\
             &      &       & 1.938  & $\mathrm{yz}$ &  22.2    \\
             &      &       & 1.972  & $\mathrm{yz}$ & -29.7    \\
\emph{alt}   & 2D   & 145   & 0.720  & $\mathrm{yz}$ & -40.2    \\
             &      &       & 0.912  & $\mathrm{yz}$ & -32.9    \\
 CdSe        & bulk & 90    & 0.906  & $\mathrm{zz}$ & -26.9    \\
 GaAs        & bulk & 90    & 2.312  & $\mathrm{xx}$ & -21.6    \\
\hline
\end{tabular}

@@

    \caption{Comparison of most intense responses of
    $\mathcal{V}^{\mathrm{ab}}$ for 2D \emph{alt} and \emph{up}, and bulk CdSe
    and GaAs structures and their corresponding polarization angle $\alpha$.}
    \label{fig:vab-str-comp}
\end{figure}

Using  Eq. \eqref{eq:vab-aw}, we calculated 
$\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)$ for the \emph{up} and
\emph{alt} 2D structures and for CdSe and GaAs which are bulk systems; 
the results
are presented in Fig. \ref{fig:vab-str-comp}. 
The angle $\alpha$ presented in
the response of each structure is that for which the response is maximized in
each case.
% 
The most intense response corresponds to the \emph{up} structure centered at
0.088\,eV 
corresponding to the long-wavelength infrared radiation (LWIR) 
and reaching a $\calv^{\rmy\rmz}=87.2$ Km/s, which means the
electrons moving along $y$ (parallel to the surface) with the spin
polarized along $z$ (perpendicular to the surface),
with a speed of 87.2 Km/s.
% 
For an energy range from 0.66\,eV to 3.0\,eV, 
corresponding
to energies of the Near Infrared (NIR) to visible radiation (VIS), 
all
four systems have contributions of the same order of magnitude.
% 
For this energy range the \emph{up} structure has two peaks centered at
1.938\,eV and 1.972\,eV reaching $\calv^{\rmy\rmz}=22.2$ Km/s and
$\calv^{\rmy\rmz}=-29.7$ Km/s and the 
\emph{alt} structure has two peaks centered at 0.720\,eV
and 0.912\,eV reaching 
$\calv^{\rmy\rmz}=-40.2$ Km/s
and
$\calv^{\rmy\rmz}=-32.9$ Km/s, 
respectively.
% 
Then, for the bulk structures we have that the CdSe has only one intense
response centered at 0.906\,eV reaching 
$\calv^{\rmz\rmz}=-26.9$ Km/s, 
and the
GaAs structure response reaches the maximum centered at
2.312\,eV reaching
$\calv^{\rmx\rmx}=\calv^{\rmy\rmy}=-21.6$ Km/s.
% 
In table \ref{tab:vab-str-comp} we present the comparison of these values for
the 2D and bulk structures; a positive (negative) spin-velocity means that the
velocity moves (anti-)parallel to the electric field.
% 
We remark that the 2D structures have resonances that give 
larger values for $\calv^{\rma\rmb}$ than the
bulk crystals; in particular
the \emph{up} structure is
$\sim 5$ times more intense.

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=\linewidth]{upplots/up-vsz-3d-1}
    \\
    \includegraphics[width=\linewidth]{upplots/up-3d-svaz-2}

    \caption{$\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega,\alpha)$ response
    as a function of the photon energy and polarization angle $\alpha$ for the
    \emph{up} structure for two energy ranges. The absolute maxima is located
    for an energy range from 0.08\,eV to 0.10\,eV, in the LWIR
   range, and two local maxima from 1.90\,eV to 1.93\,eV and from
    1.96\,eV to 2.0\,eV, in the VIS range, all for polarization
    angles between $25^{\circ}$ and $50^{\circ}$.}
    \label{fig:up-vsz-3d}   
\end{figure}

\begin{figure}[t]
    \centering
    \includegraphics[width=\linewidth]{upplots/up-vaz-rag-1}
    \\
    \vspace{4mm}
    \includegraphics[width=\linewidth]{upplots/up-vaz-rag-2}

    \caption{For the {\it up} structure and  fixing the energy to
      0.088 eV (top panel) and 1.972 eV (lower panel), 
  we show 
    $\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega,\alpha)$ (left
    scale), the corresponding velocity angle
    $\gamma_{\mathrm{z}}(\omega,\alpha)$ (right scale), the
    collinear (green arrow) and perpendicular (blue arrow) angles.
}
    \label{fig:up-vaz-rag}
\end{figure}

@@

Using Eq. \eqref{eq:vs-mag}, we calculated 
$\mathcal{V}_{\sigma^{\mathrm{b}}}(\omega,\alpha)$ 
and made the analysis for the case when the spin is fixed
along $z$, i.e. directed perpendicularly to the surface of the \emph{up} and
\emph{alt} structures. Also, using Eq. \eqref{eq:gamma-ang}, we determined the
angle $\gamma_{\mathrm{b}}(\omega,\alpha)$ where the spin-velocity is directed
along the surface of the each structure.
%

\subsubsection{{\it up} structure}\label{up:fs}

We present in Fig. \ref{fig:up-vsz-3d} 
$\mathcal{V}_{\sigma^{\mathrm{z}}} (\omega,\alpha)$
% (Eq. \eqref{eq:vs-mag})
for the two energy ranges of the \emph{up} structure that give the
highest values for $\calv^{\rmy\rmz}$ shown in Fig.~\ref{fig:vab-str-comp}.
% 
We obtain that the zone where the maximum response is reached corresponds to
a energy range of the incident beam from 0.084\,eV to 0.093\,eV, 
corresponding to the LWIR,
and polarization angles $\alpha$ 
between $30^{\circ}$ and $45^{\circ}$. 
Also,  two local maxima are held for same 
$\ga$ between $30^{\circ}$ and $45^{\circ}$, 
but for an energy range between 1.90\,eV and 2.05\,eV, in the VIS
range.
% %%%%%%%%%%%%%%%%% 
% %% 0.0 eV - 0.16 eV
% %% Description of UP |V_{s^z}| & gamma  vs. alpha agle
% %%%%%%%%%%%%%%%%%

In the top panel of Fig. \ref{fig:up-vaz-rag} we first analyze the
response for
$\hbar\go= 0.088$ eV which gives the maximum as shown
in Fig. \ref{fig:up-vsz-3d} . 
We show 
$\mathcal{V}_{\sigma^{\mathrm{z}}} (\omega,\alpha)$
as a function of $\ga$
(left scale, black solid line). 
The absolute maximum is obtained when
$\alpha = 40^{\circ}$ resulting in a value of
$\mathcal{V}_{\sigma^{\mathrm{z}}} (\omega,\alpha) = 95.8$\,Km/s, 
that according to Eq.~\eqref{eq:vs-mag} 
comes from 
$\mathcal{V}^{\mathrm{xz}}(\omega,\alpha)=39.8$\,Km/s and
$\mathcal{V}^{\mathrm{yz}}(\omega,\alpha)=87.2$\,Km/s.
% 
In the same panel we present the corresponding velocity angle
$\gamma_{\mathrm{z}}(\omega,\alpha)$ (right scale, red dashed line), that for
the absolute maximum at $\ga=40^\circ$ is
$\gamma_{\mathrm{z}}(\hbar\omega=0.088\text{eV},\alpha=40^\circ) = 65^{\circ}$. 
This results means that at $\hbar\go=0.088$ eV and a polarization
angle of the incoming 
electric field of $\ga=40^\circ$ 
the electrons with a spin pointing along $z$ move at angle of 
$\gamma_{\mathrm{z}}(\hbar\omega=0.088\text{eV},\alpha=40^\circ) =
65^{\circ}$
with respect to the $x$ direction,  
with a  maximum speed
$\mathcal{V}_{\sigma^{\mathrm{z}}} (\omega,\alpha)$ of  95.8 Km/s.
Also,
from Eq. \eqref{eq:gamma-par} we find that 
$\gamma^\parallel_{\mathrm{z}}(\omega,\alpha)=\ga=65.4^\circ$, with
$\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega,\alpha) = 82.3$ Km/s 
(see green arrow),  
and 
that from Eq. \eqref{eq:gamma-perp}, 
$\gamma^\perp_{\mathrm{z}}(\omega,\alpha)=\ga-90^\circ=65.5^\circ$,
gives $\ga=155.5^\circ$, with  
$\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega,\alpha) = 24.8$ Km/s
(see blue arrow); thus, for $\hbar\go=0.088$ eV, an
$\alpha=65.4^\circ$ ($\alpha=155.5^\circ$) gives 
electrons with spin polarized along $z$
moving 
parallel (perpendicular) to the
incident electric field,  with a speed of 82.3 (24.8) Km/s. 
% 
We also find that for all values of $\ga$,
$\gamma_{\mathrm{z}}(\omega,\alpha) =65.4^{\circ}\pm 0.3^{\circ}$,
being almost constant, and 20 Km/s $<\calv_{\gs^z}(\go,\ga)<$ 95.8 Km/s.
%

In the bottom panel of Fig. \ref{fig:up-vaz-rag}, now we take
$\hbar\go=1.972$ eV which is the local maximum 
shown in the bottom panel of Fig. \ref{fig:up-vsz-3d}. 
The local maximum is
again at  $\alpha = 40^{\circ}$ with
$\mathcal{V}_{\sigma^{\mathrm{z}}} (\omega,\alpha) =
30.3$ Km/s and $\gamma_{\mathrm{z}}(\omega,\alpha) =
78.2^{\circ}$.
% 
In this case,  
$\alpha =\gamma_{\mathrm{z}}^\parallel(\omega,\alpha) = 78.4^{\circ}$ with
$\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega,\alpha) =18.8$ Km/s 
(green circled box and arrow),
and
% 
$\gamma_{\mathrm{z}}^\perp(\omega,\alpha) = 78.5^{\circ}$, for 
$\alpha =168.5^ {\circ}$ 
(blue squared box and arrow).
% 
We find  that the velocity angle is almost constant at
$78.4^{\circ}\pm 0.3^\circ$ 
for  $0^{\circ} \leq
\alpha \leq 100^{\circ}$,
and 1 Km/s $<\calv_{\gs^z}(\go,\ga)<$ 30.3 Km/s.
% 
Although no figures are presented, we also made the analysis for the
cases when the spin polarization is directed 
along $\mathrm{x}$ and $\mathrm{y}$, finding 
the maxima for
$\hbar\go=0.088$ eV and $\alpha=40^{\circ}$, with
$\mathcal{V}_{\sigma^{\mathrm{x}}}(\omega,\alpha)=37.4$ Km/s and
$\mathcal{V}_{\sigma^{\mathrm{y}}}(\omega,\alpha)=24.8$ Km/s,
$\gamma_{\mathrm{x}}(\omega,\alpha) = 12.5^{\circ}$, and 
$\gamma_{\mathrm{y}}(\omega,\alpha) = 159.7^{\circ}$

@@

\begin{figure}[tb]
    \centering
    \includegraphics[width=\linewidth]{altplots/alt-3d-svaz}
    \caption{$|\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega,\alpha)|$ response
    as a function of the photon energy and polarization angle $\alpha$  for the
    \emph{alt} structure. The local and the absolute maxima are located in the
    energy ranges from 0.67\,eV to 0.73\,eV and from 0.90\,eV to 0.93\,eV,
    respectively, and both in the Near Infrared and for polarization angles
    between $120^{\circ}$ and $150^{\circ}$.}
    \label{fig:alt-3d-vsb}
\end{figure}

\begin{figure}[t]
    \centering
    \includegraphics[width=\linewidth]{altplots/alt-vaz-rag}
    \caption{Most intense response of
    $|\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega,\alpha)|$ (top frame, left
    scale) the corresponding velocity angle $\gamma_{\mathrm{z}}(\omega)$ (top
    frame, right scale), the collinear (circled box) and perpendicular (square
    box) angles, and the two components $\mathcal{V}^{\mathrm{xz}}(\omega)$ and
    $\mathcal{V}^{\mathrm{yz}}(\omega)$ (bottom frame) for the \emph{alt}
    structure fixing the energy to 0.912\,eV.}
    \label{fig:alt-vaz-rag}
\end{figure}

@@

\subsubsection{{\it alt} structure}

>>>>

Graphene, an allotrope of carbon with hexagonal 2D lattice structure presents
properties like fractional quantum Hall effect at room temperature, excellent
thermal transport properties, excellent conductivity\cite{heerscheNat07} and
strength \cite{geimNM07, reinaNL08, novoselov2S07, balandinNL08} being then a
perfect platform to be used in two-dimensions electronic systems; however most
electronic applications are disabled by the absence of a semiconducting gap.
Recent studies demonstrate that the band gap of graphene can be opened by
applying an electric field,\cite{zhangN09} reducing the surface
area,\cite{hanPRL07} or applying uniaxial strain.\cite{niACSN08} Another
possibility to open the gap is by doping; this has been successfully achieved
using nitrogen,\cite{weiNL2009} boron-nitrogen,\cite{guoIJ11}
silicon,\cite{colettiPRB10} noble-metals,\cite{varykhalovPRB10} and
hydrogen.\cite{eliasS09, guisingerNL09, samarakoonACSN10}
% 
Depending on the percentage of hydrogenation and spatial configurations of
hydrogen-carbon bonds, hydrogenated graphene can result in different spatial
configurations.
% 
In this paper we present two 50\% hydrogenated graphene noncentrosymmetric
structures both presenting a discernible band gap: the \emph{Up} structure,
shown in Fig. \ref{fig:up-struc}, has hydrogen atoms bonded to the carbon layer
only in the upper side of the structure while the \emph{Alt} structure, shown
in Fig. \ref{fig:alt-struc}, has hydrogen alternating in the upper and bottom
sides of the carbon slab.\cite{zapataPSB2016}

@@

This paper is organized as follows. In Section \ref{sec:theory} we present the
theory and formulas that describe PSC and SVI. In Section \ref{sec:results} we
describe the details of calculations and the corresponding SVI spectra for the
\emph{Up} and \emph{Alt} structures. Finally, we present our conclusions in
Section \ref{sec:conclusions}.

@@

\caption{Unit cell of \emph{Up} structure. Layer division, atom types and
positions for the \emph{Up} structure. The structure unit cell was divided in
two layers corresponding to hydrogen and carbon atoms.The corresponding layer
atom position is depicted in Fig. \ref{fig:up-struc} with the corresponding
number of layer.}
\label{tab:up-unitcell}
\end{table}
% 
% 
\begin{table}[t]
\center
\begin{tabular}{ccccc}\\
\hline
\quad Layer \quad & \quad Atom \qquad & \multicolumn{3}{c}{Position [\AA]} \\
\cline{3-5}
\quad No.   \quad & \quad type \qquad & $x$ & $y$ & $z$  \\
\hline
1 & H &  -0.61516 &  -1.42140 & \ 1.47237 \\
2 & C &  -0.61516 &  -1.73300 & \ 0.39631 \\
3 & C & \ 0.61516 & \ 1.73300 & \ 0.15807 \\
4 & C & \ 0.61516 & \ 0.42201 &  -0.15814 \\
5 & C &  -0.61516 &  -0.37396 &  -0.39632 \\
6 & H &  -0.61516 &  -0.68566 &  -1.47237 \\
\hline
\end{tabular}

\caption{Unit cell of \emph{Alt} structure. Layer division, atom types and
positions for the \emph{Alt} structure. The structure unit cell was divided in
six layers corresponding each one to atoms in different $z$ positions. The
corresponding layer atom position is depicted in Fig. \ref{fig:alt-struc} with
the corresponding number of layer.}
\label{tab:alt-unitcell}
\end{table}

We preset the results for 
and
$\mathcal{V}_{\sigma^{\mathrm{b}}}(\omega,\alpha)$ 
and
$\mathcal{V}_{\mathrm{a}}(\omega,\alpha)$ 
for the
C$_{16}$H$_{8}$-up  and C$_{16}$H$_{8}$-alt 
structures being both
noncentrosymmetric semi-infinite 2D carbon systems with 50\% hydrogenation in
different arrangements. We recall that the \emph{Up} structure has hydrogen
atoms only on the upper side of the carbon sheet while the \emph{Alt} structure
has alternating hydrogen atoms on the upper and bottom sides. Also we take the
hexagonal carbon lattice to be on the $xy$ plane for both structures, and the
carbon-hydrogen bonds on the perpendicular $xz$ plane, as depicted in Figs.
\ref{fig:up-struc} and \ref{fig:alt-struc}. The coordinates for the
\emph{Up} and \emph{Alt} unit cells of the structures are presented in Tables
\ref{tab:up-unitcell} and \ref{tab:alt-unitcell}.

We calculated the self-consistent ground state and the Kohn-Sham
states using density functional theory in the local density approximation (DFT-
LDA) with a planewave basis using the ABINIT code \cite{gonzeCPC09}.
% 
We used Hartwigsen-Goedecker-Hutter (HGH) relativistic separable dual-space
Gaussian pseudopotentials \cite{hartwigsenPRB98} including the spin-orbit
interaction needed to calculate $\mu^{\mathrm{abcd}}(\omega,\alpha)$ presented
in Eq. \eqref{eq:mu}.
% 
The convergence parameters for the calculations of our results corresponding to
the \emph{Up} and \emph{Alt} structures are cutoff energies of 40\,Ha and
65\,Ha, resulting in LDA energy band gaps of 0.084\,eV and 0.718\,eV,
respectively. The energy eigenvalues and matrix elements for the \emph{Up} and
\emph{Alt} structures were calculated using 12802 $\mathbf{k}$ points and 14452
$\mathbf{k}$ points in the IBZ to integrate $\mu^{\mathrm{abcd}}(\go)$ and
$\xi^{\rma\rmb}(\go)$ using the linearized analytic tetrahedron method
(LATM).\cite{nastosPRB07} We neglect the anomalous velocity term
$\hbar(\bfgs\times\nabla V)/4m^2c^2$, where $V$ is the crystal potential, in
$\hat\bfv$ of Eq.~\eqref{z.1}, as this term is known to give small contribution
to PSC.\cite{bhatPRL05} Therefore, $[\hat\bfv,\hat\bfS]=0$ and Eq.~\eqref{z.1}
reduces to $\hat K^{\rma\rmb}=\hat v^\rma \hat S^\rmb=\hat S^\rmb \hat v^\rma$.

@@

\begin{table}[b]
\begin{tabular}{cccccc}
\hline
\multirow{2}{*}{Structure \quad} & 
Kind of \quad & 
Pol. &
Energy & 
\multicolumn{2}{c}{$\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)$}\\
\cline{5-6}
& system & Ang. & [eV] & $\mathrm{ab}$ \quad & [Km/s]\\
\hline
\emph{Up}    & 2D   & 35    & 0.084  & $\mathrm{yz}$ &  684.2    \\
             &      &       & 1.954  & $\mathrm{yz}$ &  293.9    \\
             &      &       & 1.958  & $\mathrm{yz}$ & -273.0    \\
\emph{Alt}   & 2D   & 150   & 0.720  & $\mathrm{yz}$ & -722.2    \\
             &      &       & 0.911  & $\mathrm{yz}$ & -357.8    \\
 CdSe        & bulk & 90    & 0.912  & $\mathrm{zz}$ &  -32.3    \\
 GaAs        & bulk & 90    & 2.326  & $\mathrm{xx}$ &  -29.9    \\
\hline
\end{tabular}

@@

    \caption{Comparison of most intense responses of
    $\mathcal{V}^{\mathrm{ab}}$ for 2D \emph{Alt} and \emph{Up}, and bulk CdSe
    and GaAs structures and their corresponding polarization angle $\alpha$.}
    \label{fig:vab-str-comp}
\end{figure}

Using  Eq. \eqref{eq:vab-aw}, we calculated 
$\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)$ for the \emph{Up} and
\emph{Alt} 2D structures and for CdSe and GaAs which are bulk systems; 
the results
are presented in Fig. \ref{fig:vab-str-comp}. 
The angle $\alpha$ presented in
the response of each structure is that for which the response is maximized in
each case.
% 
The most intense response corresponds to the \emph{Alt} structure centered at
0.720\,eV corresponding to the short-wavelength infrared radiation (SWIR) and
reaching a $\calv^{\rmy\rmz}=-722.2$ Km/s, which means the electrons moving
along $y$ (parallel to the surface) with the spin polarized along $z$
(perpendicular to the surface), with a speed of -722.2 Km/s.
% 
For an energy range from 0.66\,eV to 3.0\,eV, corresponding to energies of the
Near Infrared (NIR) to visible radiation (VIS), the \emph{Up} structure has two
peaks centered at 1.954\,eV and 1.958\,eV reaching $\calv^{\rmy\rmz}=293.9$
Km/s and $\calv^{\rmy\rmz}=-273.0$ Km/s and the \emph{Alt} structure one more
peak centered at 0.0.911\,eV reaching $\calv^{\rmy\rmz}=-357.83$ Km/s.
% 
Then, for the bulk structures we have that the CdSe has only one intense
response centered at 0.912\,eV reaching $\calv^{\rmz\rmz}=-32.3$ Km/s, and the
GaAs structure response reaches the maximum centered at 2.326\,eV reaching
$\calv^{\rmx\rmx}=\calv^{\rmy\rmy}=--28.9$ Km/s.
% 
In table \ref{tab:vab-str-comp} we present the comparison of these values for
the 2D and bulk structures; a positive (negative) spin-velocity means that the
velocity moves (anti-)parallel to the electric field.
% 
We remark that the 2D structures have resonances that give larger values for
$\calv^{\rma\rmb}$ than the bulk crystals; in particular the \emph{Alt}
structure is $\sim 22$ times more intense.

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=\linewidth]{upplots/up-vsz-w1}
    \caption{Caption here}
    \label{fig:up-vsz-w1}
\end{figure}

\begin{figure}[t]
    \centering
    \includegraphics[width=\linewidth]{upplots/up-vsz-w2}
    \caption{caption here}
    \label{fig:up-vsz-w2}
\end{figure}

@@

Using Eq. \eqref{eq:vs-mag}, we calculated $\mathcal{V}_{\sigma^{\mathrm{b}}}
(\omega,\alpha)$ and made the analysis for the case when the spin is fixed
along $z$, i.e. directed perpendicularly to the surface of the \emph{Up} and
\emph{Alt} structures. Also, using Eq. \eqref{eq:gamma-ang}, we determined the
angle $\gamma_{\mathrm{b}}(\omega,\alpha)$ where the spin-velocity is directed
along the surface of the each structure.
%

\subsubsection{\emph{Up} structure}\label{up:fs}

We present in Figs. \ref{fig:up-vsz-w1} and \ref{fig:up-vsz-w2}
$\mathcal{V}_{\sigma^{\mathrm{z}}} (\omega,\alpha)$ for the two energy ranges
of the \emph{Up} structure that give the highest values for $\calv^{\rmy\rmz}$
shown in Fig.~\ref{fig:vab-str-comp}.
% 
We obtain that the zone where the maximum response is reached corresponds to a
energy range of the incident beam from 0.084\,eV to 0.093\,eV, corresponding to
the LWIR, and polarization angles $\alpha$ between $35^{\circ}$ and
$55^{\circ}$. Also,  two local maxima are held for same $\ga$ between
$35^{\circ}$ and $55^{\circ}$, but for an energy range between 1.950\,eV and
1.960\,eV, in the VIS range.
% %%%%%%%%%%%%%%%%% 
% %% 0.0 eV - 0.16 eV
% %% Description of UP |V_{s^z}| & gamma  vs. alpha agle
% %%%%%%%%%%%%%%%%%

In bottom panel of Fig. \ref{fig:up-vsz-w1} we first analyze the response for
$\hbar\go= 0.084$ eV which gives the maximum. We show
$\mathcal{V}_{\sigma^{\mathrm{z}}} (\omega,\alpha)$ as a function of $\ga$
(left scale, black solid line). The absolute maximum is obtained when $\alpha =
35^{\circ}$ resulting in a value of $\mathcal{V}_{\sigma^{\mathrm{z}}}
(\omega,\alpha) = 758.2$\,Km/s, that according to Eq.~\eqref{eq:vs-mag} comes
from $\mathcal{V}^{\mathrm{xz}}(\omega,\alpha) = 326.7$\,Km/s and
$\mathcal{V}^{\mathrm{yz}}(\omega,\alpha) = 684.2$\,Km/s.
% 
In the same panel we present the corresponding velocity angle
$\gamma_{\mathrm{z}}(\omega,\alpha)$ (right scale, red line), that for the
absolute maximum at $\ga = 35^\circ$ is $\gamma_{\mathrm{z}}(\hbar\omega =
0.084\text{eV},\alpha = 35^\circ) = 64.475^{\circ}$. This results means that at
$\hbar\go=0.084$ eV and a polarization angle of the incoming electric field of
$\ga = 35^\circ$ the electrons with a spin pointing along $z$ move at angle of
$\gamma_{\mathrm{z}}(\hbar\omega = 0.084\text{eV},\alpha=35^\circ) \sim
64.5^{\circ}$ with respect to the $x$ direction, with a  maximum speed
$\mathcal{V}_{\sigma^{\mathrm{z}}} (\omega,\alpha)$ of  758.2 Km/s. Also, from
Eq. \eqref{eq:gamma-par} we find that
$\gamma^\parallel_{\mathrm{z}}(\omega,\alpha)=\ga=64.475^\circ$, with
$\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega,\alpha) = 646.0$ Km/s (see green
arrow), and that from Eq. \eqref{eq:gamma-perp},
$\gamma^\perp_{\mathrm{z}}(\omega,\alpha)=\ga-90^\circ=64.47^\circ$, gives
$\ga=154.47^\circ$, with $\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega,\alpha) =
195.6$ Km/s (see blue arrow); thus, for $\hbar\go=0.084$ eV, an $\alpha \sim
65.47^\circ$ ($\alpha \sim 154.5^\circ$) gives electrons with spin polarized
along $z$ moving parallel (perpendicular) to the incident electric field,  with
a speed of 646.0 (195.6) Km/s.
% 
We also find that for all values of $\ga$, $\gamma_{\mathrm{z}}(\omega,\alpha)
\sim 64.48^{\circ}$, being almost constant, and 180.0 Km/s
$<\calv_{\gs^z}(\go,\ga)<$ 758.2 Km/s.

In the bottom panel of Fig. \ref{fig:up-vsz-w2}, now we take
$\hbar\go=1.954$ eV which is the local maximum 
shown in the to panel of same figure. 
The local maximum is
again at  $\alpha = 35^{\circ}$ with
$\mathcal{V}_{\sigma^{\mathrm{z}}} (\omega,\alpha) =
300.4$ Km/s and $\gamma_{\mathrm{z}}(\omega,\alpha) =
78.0^{\circ}$.
% 
In this case, $\alpha =\gamma_{\mathrm{z}}^\parallel(\omega,\alpha) =
78.0^{\circ}$ with $\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega,\alpha) =183.7$
Km/s (green circled box and arrow), and
$\gamma_{\mathrm{z}}^\perp(\omega,\alpha) = 78.0^{\circ}$, for $\alpha =168.0^
{\circ}$ (blue squared box and arrow).
% 
We find  that the velocity angle is almost constant having a range from
$77^{\circ}$ to $78.04^{\circ}$ for  $0^{\circ} \leq
\alpha \leq 180^{\circ}$, and 6.8 Km/s $<\calv_{\gs^z}(\go,\ga)<$ 300.4 Km/s.
% 
Although no figures are presented, we also made the analysis for the
cases when the spin polarization is directed 
along $\mathrm{x}$ and $\mathrm{y}$, finding 
the maxima for
$\hbar\go=0.084$ eV and $\alpha=35^{\circ}$, with
$\mathcal{V}_{\sigma^{\mathrm{x}}}(\omega,\alpha)=219.6$ Km/s and
$\mathcal{V}_{\sigma^{\mathrm{y}}}(\omega,\alpha)=280.3$ Km/s,
$\gamma_{\mathrm{x}}(\omega,\alpha) = 79.3^{\circ}$, and 
$\gamma_{\mathrm{y}}(\omega,\alpha) = 78.2^{\circ}$.

@@

\begin{figure}[tb]
    \centering
    \includegraphics[width=\linewidth]{altplots/alt-vsz}
    \caption{Caption here}
    \label{fig:alt-vsz}
\end{figure}

@@

\subsubsection{\emph{Alt} structure}

We proceed to analyze the $\emph{Alt}$ structure, just as we did the \emph{Up}
structure. Fig. \ref{fig:alt-vsz} shows $\mathcal{V}_{\sigma^{\mathrm{z}}}
(\omega,\alpha)$ (Eq. \eqref{eq:vs-mag}), where we see that the maximum
response is reached for $\hbar\go=0.720$ eV and a local maximum is obtained at
$\hbar\go=0.912$ eV, both located in the short-wavelength infrared radiation
(SWIR) range and for polarization angles, $\ga$,  between $135^{\circ}$ and
$160^{\circ}$.
